A Derivative Tracking Model for Wind Power Forecast Error

### 2020-07-11 LaTeX_Code/Paper/aoas-template.tex
Added latest paper version

\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{bm}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{todonotes}
%\usepackage[normalem]{ulem}
\usepackage{float}
\usepackage{commath}
\usepackage{fancybox}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{url}
\usepackage[margin=0.5in]{geometry}
\usepackage[backend=biber, style=authoryear, citestyle=authoryear, autocite=inline, isbn=true]{biblatex} % package for bibliography 
%\DeclareLanguageMapping{english}{english-apa}
\definecolor{lime}{HTML}{A6CE39}

\DeclareRobustCommand{\orcidicon}{
	\begin{tikzpicture}
	\draw[lime, fill=lime] (0,0) 
	circle [radius=0.16] 
	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
	\draw[white, fill=white] (-0.0625,0.095) 
	circle [radius=0.007];
	\end{tikzpicture}
	\hspace{-2mm}
}
\foreach \x in {A, ..., Z}{\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}
			{\noexpand\orcidicon}}
}

\newcommand{\orcidauthorA}{0000-0003-3220-0923} % For author RC
\newcommand{\orcidauthorB}{0000-0003-1967-4446} % For author RT
\newcommand{\orcidauthorC}{0000-0003-3468-8811} % For author AK
\newcommand{\orcidauthorD}{0000-0001-5114-853X} % For author MS

\theoremstyle{definition}
\newtheorem{Def}{Definition}[section]
\newtheorem{Eg}{Example}[section]
\newtheorem{Rm}{Remark}[section]
\newtheorem{Ex}{Exercise}[section]
\newtheorem{Prop}[Def]{Proposition}
\newtheorem{Lem}[Def]{Lemma}
\newtheorem{Thm}[Def]{Theorem}
\newtheorem{Cor}[Def]{Corollary}
\newtheorem{Rem}[Def]{Remark}
\newtheorem{Ass}[Def]{Assumption}
\newtheorem{Met}[Def]{Method}
\newtheorem{Res}[Def]{Result}
\newcommand{\red}{\textcolor{black}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}

\addbibresource{A_Derivative_Tracking_Model_for_Wind_Power_Forecast_Error.bib}

%\renewbibmacro*{title}{%
%  \ifboolexpr{
%    test {\iffieldundef{title}}
%    and
%    test {\iffieldundef{subtitle}}
%  }
%    {}
%    {\printtext{%
%     \printtext[titlecase]{\usefield{\uline}{title}}%
%     \setunit{\subtitlepunct}%
%     \printfield[titlecase]{subtitle}}%
%     \newunit}%
%  \printfield{titleaddon}}

\setcounter{page}{1}

@@

\title{A Derivative Tracking Model for Wind Power Forecast Error}  % \\  \small{Report}}

%\author[1]{Waled Alhaddad}
\author[1]{Renzo Caballero\orcidA{}}
\author[2]{Ahmed Kebaier\orcidC{}}
\author[3]{Marco Scavino\orcidD{}}
\author[4]{Ra\'ul Tempone\orcidB{}}
%\author[2]{Corresponding Author\thanks{email@2nduniversity.com}}
\affil[1,4]{CEMSE Division, King Abdullah University of Science and Technology (KAUST), Saudi Arabia}
\affil[2]{\red{Universit\'e Sorbonne Paris Nord, LAGA, CNRS, UMR 7539, F-93430, Villetaneuse, France}}
\affil[3]{Universidad de la Rep\'ublica, Instituto de Estad\'{\i}stica (IESTA), Montevideo, Uruguay}
\affil[4]{Alexander von Humboldt Professor, RWTH Aachen University, Germany}
\date{}

%\author{ Waled Alhaddad \textsuperscript{\textasteriskcentered} \qquad Ahmed Kebaier\textsuperscript{\ddag} \qquad Ra\'ul  Tempone\textsuperscript{\textasteriskcentered}\textsuperscript{\textdagger} \\
%\textsuperscript{\textasteriskcentered}CEMSE Division, King Abdullah University of Science and Technology (KAUST), Saudi Arabia \\ \textsuperscript{\textdagger}Alexander von Humboldt Professor, RWTH Aachen University, Germany
% \\ \textsuperscript{\ddag}Université Paris 13, Sorbonne Paris Cité, LAGA, CNRS (UMR 7539), Villetaneuse, France}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT

\begin{abstract}

Reliable wind power generation forecasting is crucial for applications such as the allocation of energy reserves, optimization for electricity price, and operation scheduling of conventional power plants. We propose a data-driven model based on parametric Stochastic Differential Equations (SDEs) to capture the real asymmetric dynamics of wind power forecast errors. Our SDE framework features time-derivative tracking of the forecast, time-varying mean-reversion parameter, and an improved state-dependent diffusion term. The methodology we developed allows the simulation of future wind power production paths and to obtain sharp empirical confidence bands. All the procedures are agnostic of the forecasting technology, and they enable comparisons between different forecast providers. We apply the model to historical Uruguayan wind power production data and forecasts between April and December 2019.

\end{abstract}

Keywords: wind power, probabilistic forecasting, stochastic differential equations, Lamperti transform, numerical optimization, model selection, time-inhomogeneous Jacobi diffusion

2020 AMS Subject Classification: 60H10, 62M20, 65K10.

%%%%%%%%% BODY TEXT

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/four_days/377.eps}
\includegraphics[width=0.45\textwidth]{plots/four_days/417.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{plots/four_days/437.eps}
\includegraphics[width=0.45\textwidth]{plots/four_days/477.eps}
\caption{Four 24-hour segments with the wind power real production in Uruguay (orange line) recorded every ten minutes, and the hourly wind power production forecasted by provider A (blue line).}
  \label{fig:sample_data}
\end{figure}

@@

Let $\bm{\theta} = (\theta_0,\alpha)$, and choose a state-dependent diffusion term that avoids the process exiting from the range $[0,1]$ as follows:
  \begin{equation}
    b (X_t; \bm{\theta} )= \sqrt{2 \alpha \theta_0 X_t (1 - X_t)}
  \end{equation}
  where $\alpha >0$ is an unknown parameter that controls the path variability. This diffusion term belongs to the Pearson diffusion family and, in particular, it defines a Jacobi type diffusion. It is useful to recall that (\cite[440]{foso}) a Pearson diffusion is a stationary solution to a stochastic differential equation of the form \begin{equation}
    \dif X_t = - \theta (X_t - \mu)\dif t + \sqrt{2 \theta \left(a X_t^2 + b X_t + c\right)} \dif W_t
    \label{eq:poly_diff}
  \end{equation}
where $\theta>0$, and $a$, $b$, and $c$ are parameters such that the square root is well defined when $X_t$ is in the state space. These parameters, together with $\mu$, the mean of the invariant distribution, determine the state space of the diffusion as well as the shape of the invariant distribution.

An exhaustive classification of the (stationary) Pearson diffusions is presented in (\cite[440-443]{foso}) where, in particular, it is discussed the case $a < 0$ and $b(x; \bm{\theta}) = \sqrt{2 a \theta x (x-1)}$, where the invariant distribution is a Beta distribution with parameters $\left( \frac{\mu}{-a}, \frac{1 - \mu}{-a} \right),$ that leads to the well-known Jacobi diffusions, so-called because the eigenfunctions of the infinitesimal generator of these processes are the Jacobi polynomials (see, for example, \cite[2860-2861]{leph}).

@@

We consider the following Lamperti transform with unknown parameters
%{\color{red} (Remind that continuous-time models with different diffusion term are singular and the likelihood cannot be written down. Besides, it can be unstable to work with low frequency. If the frequency of the data is too low, the Lamperti transform helps to detect this problem (in principle, it is not clear how to determine the right frequency for estimation. )}
%\begin{equation}
%Z_t = h(V_t, t) = \frac{1}{\sqrt{2 \theta_0 \alpha}} \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
%\end{equation}
%where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, 
\begin{align}
Z_t = h(V_t, t; \bm{\theta} )  & = \frac{1}{\sqrt{2 \alpha \theta_0 }} \int \frac{1}{\sqrt{(v + p_t)(1 - v - p_t)}} \dif v \Bigg{\vert}_{v = V_t}   = - \sqrt{ \frac{2}{ \alpha \theta_0 }} \arcsin (\sqrt{ 1 - V_t - p_t}) \label{eq:LampZ}
\end{align}
that, after applying It\^{o}'s formula on $h(V_t, t; \bm{\theta} )$, leads to the following SDE with state independent unit diffusion term
\begin{align}
\dif Z_t = \Bigg[  & \frac{\dot{p}_t}{ \sqrt{2 \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}}  + \frac{- \theta_t V_t}{ \sqrt{2  \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{4} \frac{\sqrt{2 \alpha \theta_0} \left( 1 - 2 (V_t + p_t)\right)}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}}  \Bigg] \dif t + \dif W_t \,. \label{eq:stindepSDE}
\end{align}
After replacing $V_t = 1 - p_t - \sin^2 \left(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \right) $ in (\ref{eq:stindepSDE}), we obtain that the  process $Z_t$ satisfies the SDE
\begin{align}
\dif Z_t & = \left[ \frac{ \dot{p}_t  - \theta_t  \left(1 - p_t - \sin^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\sqrt{2 \alpha \theta_0} \cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)}   - \frac{1}{4}  \frac{\sqrt{2 \alpha \theta_0} \left(1 - 2 \cos^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)} \right]\dif t + \dif W_t \nonumber \\
&  = \left[  \frac{  2  \dot{p}_t - \theta_t (1 - 2 p_t)  + (\alpha \theta_0 - \theta_t) \cos(- \sqrt{2 \alpha \theta_0 } Z_t) }{\sqrt{2 \alpha \theta_0} \sin{(- \sqrt{2 \alpha \theta_0} Z_t)}}  \right] \dif t + \dif W_t.  \label{eq:stindepSDE2}
\end{align}

@@

\red{Remark: Moreover, this obtained Gaussian distribution supports the validity of the choice of our model diffusion coefficient given by (\ref{eq:poly_diff}). 
}
%{\color{red} Remind that Lamperti transform has the purpose tho check the consistency of the models, that is if the estimation of $\bm{\theta}$ with $V$ and $Z$ differs it is alarming.}

@@

\red{Suppose that any of $M$ non-overlapping paths of the continuous-time It\^{o} process $V = \{ V_t, t  \in [0,T] \}$, each one starting at a different time $t_j$ with $j = 1, \dots, M$, is sampled at $N + 1$ equispaced discrete points with given length interval $\Delta$. Let $ V^{M,N + 1}=\left\{ V_{t_1^{N + 1}} , V_{t_2^{N + 1}} ,\ldots , V_{t_M^{N + 1}} \right\}$ denote this random sample, with $V_{t_j^{N + 1}} =\left\{ V_{t_j + i \Delta}\,, i = 0, \ldots, N \right\}$.}

@@

However, closed-form solutions to initial-boundary value problems for time-inhomogeneous diffusions can be obtained only in a few cases (see, for example, \cite[Section 3.1]{eglix}). In our case, solving numerically (\ref{eq:fpk}) for the transition densities of the process $V$ at every transition step is computationally expensive. 
Several numerical techniques have been devised to obtain estimates for the unknown parameters of continuous-time SDE models with discrete observations (see, for example, \cite{prewo} for likelihood-based inference techniques, \cite{Sor} for an estimating function approach). As explained in the next subsection, we have considered approximate likelihood methods, similar in spirit to \autocite[Section 11.4]{saso}.

@@

Gaussian approximations to the transition densities of nonlinear time-inhomogeneous SDEs are available through different algorithms \autocite[Chapter 9]{saso}. However, as Figure \ref{fig:error_transitions} may suggest at first glance, the choice of a Gaussian density could be inadequate when straightly applied to approximate the transition density of the forecast error $V$ of the normalized wind power production.

@@

A suitable candidate for a surrogate transition density of $V$ is a Beta distribution on a compact interval parameterized by two positive shape parameters, $\xi_1, \xi_2$. \red{Recall that the choice of the beta proxy distribution is a natural choice as it is the invariant distribution of the Jacobi type processes.}

@@

The approximate log-likelihood $\tilde{\ell}(\cdot ; v^{M, N+1})$ of the observed sample $v^{M, N+1}$ can be expressed as 
\begin{equation}
 \tilde{\ell} \left(\bm{\theta}; v^{M,N +1}\right) = \sum_{j=1}^M \sum_{i=1}^N \log  \left\{ \frac{1}{2(1 - \epsilon)} \frac{1}{B(\xi_1(t_{j,i}^-), \xi_2(t_{j,i}^-))} \left( \frac{v_{j,i} + 1 - \epsilon}{2(1 - \epsilon)} \right)^{\xi_1(t_{j,i}^-) -1}  \left( \frac{1 - \epsilon - v_{j,i}}{2(1 - \epsilon)} \right)^{\xi_2(t_{j,i}^-) -1} \right\},
\label{eq:loglikelihoodV}
\end{equation}
where the shape parameters $\xi_1(t_{j,i}^-)$ and $\xi_2(t_{j,i}^-)$, according to (\ref{param_transformed_beta}), depend on the limit quantities $\mu(t_{j,i}^-;\bm{\theta} )$ and $\sigma^2(t_{j,i}^-;\bm{\theta} )$ as $t\uparrow t_{j,i}$ that are computed solving numerically the initial-value problem (\ref{Vtmom}). $B(\xi_1,\xi_2)$ denotes the Beta distribution with parameters $\xi_1$ and $\xi_2$.

@@

\red{In this subsection, we aim to infer the model's parameters using optimization techniques. We start by finding an initial guess close enough to the optimal value, and from that point, start the optimization.}

@@

Our statistical analysis starts with partitioning the 147 segments of normalized wind power production, each 24-hours long. We select 73 non-contiguous segments for the models' calibration procedure, assigning them to the training set. The other 74 non-contiguous segments compose the test set. Such an allocation mechanism guarantees independence among the segments, matching the assumption we did in Section \ref{Section_5} to formulate the statistical models. \red{Additional cross-correlation tests were performed to ensure this assumption.}

@@

As an auxiliary verification, we plot the negative log-likelihood (negative version of (\ref{eq:loglikelihoodV})) as a function of the parameters, and we use additional minimization functions from MATLAB R2019b. \red{Moreover, we realized an additional inference utilizing the test sets to guarantee the robustness of our numerical methods.}

>>>>

% Template for the submission to:
%   The Annals of Applied Statistics    [AOAS]
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% In this template, the places where you   %%
%% need to fill in your information are     %%
%% indicated by '???'.                      %%
%%                                          %%
%% Please do not use \input{...} to include %%
%% other tex files. Submit your LaTeX       %%
%% manuscript as one .tex document.         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[aoas,preprint]{imsart}

%% Packages
\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{float}
\usepackage{commath}
\usepackage{fancybox}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{url}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%\RequirePackage{graphicx}% uncomment this for including figures

\startlocaldefs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Uncomment next line to change            %%
%% the type of equation numbering           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\numberwithin{equation}{section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Axiom, Claim, Corollary, Hypothezis, %%
%% Lemma, Theorem, Proposition              %%
%% use \theoremstyle{plain}                 %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
%\newtheorem{???}{???}
%\newtheorem*{???}{???}
%\newtheorem{???}{???}[???]
%\newtheorem{???}[???]{???}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Assumption, Definition, Example,     %%
%% Notation, Property, Remark, Fact         %%
%% use \theoremstyle{remark}                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\theoremstyle{remark}
%\newtheorem{???}{???}
%\newtheorem*{???}{???}
%\newtheorem{???}{???}[???]
%\newtheorem{???}[???]{???}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please put your definitions here:        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}

\endlocaldefs

@@

\begin{frontmatter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{A Derivative Tracking Model for Wind Power Forecast Error}
%\title{A sample article title with some additional note\thanksref{T1}}
\runtitle{A Derivative Tracking Model for Wind Power Forecast Error}
%\thankstext{T1}{A sample of additional note to the title.}

\begin{aug}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Only one address is permitted per author. %%
%%Only division, organization and e-mail is %%
%%included in the address.                  %%
%%Additional information can be included in %%
%%the Acknowledgments section if necessary. %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[A]{\fnms{Renzo} \snm{Caballero}\ead[label=e1]{Renzo.CaballeroRosas@kaust.edu.sa}},
\author[B]{\fnms{Ahmed} \snm{Kebaier}\ead[label=e2,mark]{kebaier@math.univ-paris13.fr}},
\author[C]{\fnms{Marco} \snm{Scavino}\ead[label=e3,mark]{mscavino@iesta.edu.uy}}
\and
\author[D]{\fnms{Ra\'ul} \snm{Tempone}\ead[label=e4,mark]{tempone@uq.rwth-aachen.de}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Addresses                                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address[A]{CEMSE Division, King Abdullah University of Science and Technology, Saudi Arabia, \printead{e1}}
\address[B]{Universit\'e Sorbonne Paris Nord, LAGA, CNRS, UMR 7539, F-93430, Villetaneuse, France, \printead{e2}}
\address[C]{Instituto de Estad\'{\i}stica (IESTA), Universidad de la Rep\'ublica, Montevideo, Uruguay, \printead{e3}}
\address[D]{Chair of Mathematics for Uncertainty Quantification, RWTH Aachen University, Germany, \printead{e4}}
\end{aug}

\begin{abstract}
Reliable wind power generation forecasting is crucial for applications such as the allocation of energy reserves, optimization for electricity price, and operation scheduling of conventional power plants. We propose a data-driven model based on parametric Stochastic Differential Equations (SDEs) to capture the real asymmetric dynamics of wind power forecast errors. Our SDE framework features time-derivative tracking of the forecast, time-varying mean-reversion parameter, and an improved state-dependent diffusion term. The methodology we developed allows the simulation of future wind power production paths and to obtain sharp empirical confidence bands. All the procedures are agnostic of the forecasting technology, and they enable comparisons between different forecast providers. We apply the model to historical Uruguayan wind power production data and forecasts between April and December 2019.
\end{abstract}

\begin{keyword}
\kwd{Wind Power}
\kwd{Probabilistic Forecasting}
\kwd{Stochastic Differential Equations}
\kwd{Lamperti Transform}
\kwd{Numerical Optimization}
\kwd{Model Selection}
\kwd{Time-Inhomogeneous Jacobi Diffusion}
\end{keyword}

\end{frontmatter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please use \tableofcontents for articles %%
%% with 50 pages and more                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main text entry area:

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/four_days/377.eps}
\includegraphics[width=0.45\textwidth]{plots/four_days/417.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/four_days/437.eps}
%\includegraphics[width=0.45\textwidth]{plots/four_days/477.eps}
\caption{Four 24-hour segments with the wind power real production in Uruguay (orange line) recorded every ten minutes, and the hourly wind power production forecasted by provider A (blue line).}
  \label{fig:sample_data}
\end{figure}

@@

Let $\bm{\theta} = (\theta_0,\alpha)$, and choose a state-dependent diffusion term that avoids the process exiting from the range $[0,1]$ as follows:
  \begin{equation}
    b (X_t; \bm{\theta} )= \sqrt{2 \alpha \theta_0 X_t (1 - X_t)}
  \end{equation}
  where $\alpha >0$ is an unknown parameter that controls the path variability. This diffusion term belongs to the Pearson diffusion family and, in particular, it defines a Jacobi type diffusion. It is useful to recall that \cite[440]{foso} a Pearson diffusion is a stationary solution to a stochastic differential equation of the form \begin{equation}
    \dif X_t = - \theta (X_t - \mu)\dif t + \sqrt{2 \theta \left(a X_t^2 + b X_t + c\right)} \dif W_t
    \label{eq:poly_diff}
  \end{equation}
where $\theta>0$, and $a$, $b$, and $c$ are parameters such that the square root is well defined when $X_t$ is in the state space. These parameters, together with $\mu$, the mean of the invariant distribution, determine the state space of the diffusion as well as the shape of the invariant distribution.

An exhaustive classification of the (stationary) Pearson diffusions is presented in \cite[440-443]{foso} where, in particular, it is discussed the case $a < 0$ and $b(x; \bm{\theta}) = \sqrt{2 a \theta x (x-1)}$, where the invariant distribution is a Beta distribution with parameters $\left( \frac{\mu}{-a}, \frac{1 - \mu}{-a} \right),$ that leads to the well-known Jacobi diffusions, so-called because the eigenfunctions of the infinitesimal generator of these processes are the Jacobi polynomials (see, for example, \cite[2860-2861]{leph}).

@@

We consider the following Lamperti transform with unknown parameters
%{\color{red} (Remind that continuous-time models with different diffusion term are singular and the likelihood cannot be written down. Besides, it can be unstable to work with low frequency. If the frequency of the data is too low, the Lamperti transform helps to detect this problem (in principle, it is not clear how to determine the right frequency for estimation. )}
%\begin{equation}
%Z_t = h(V_t, t) = \frac{1}{\sqrt{2 \theta_0 \alpha}} \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
%\end{equation}
%where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, 
\begin{align}
\begin{split}
Z_t = h(V_t, t; \bm{\theta} ) & = \frac{1}{\sqrt{2 \alpha \theta_0 }} \int \frac{1}{\sqrt{(v + p_t)(1 - v - p_t)}} \dif v \Bigg{\vert}_{v = V_t}  \\
& = - \sqrt{ \frac{2}{ \alpha \theta_0 }} \arcsin (\sqrt{ 1 - V_t - p_t}) \label{eq:LampZ}
\end{split}
\end{align}
that, after applying It\^{o}'s formula on $h(V_t, t; \bm{\theta} )$, leads to the following SDE with state independent unit diffusion term
\begin{multline}
\begin{split}
\dif Z_t=& \Bigg[   \frac{\dot{p}_t}{ \sqrt{2 \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}}  \\
&+ \frac{- \theta_t V_t}{ \sqrt{2  \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{4} \frac{\sqrt{2 \alpha \theta_0} \left( 1 - 2 (V_t + p_t)\right)}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}}  \Bigg] \dif t + \dif W_t . \label{eq:stindepSDE}
\end{split}
\end{multline}
After replacing $V_t = 1 - p_t - \sin^2 \left(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \right) $ in (\ref{eq:stindepSDE}), we obtain that the  process $Z_t$ satisfies the SDE
{\small
\begin{align}
\begin{split}
\dif Z_t  =& \left[ \frac{ \dot{p}_t  - \theta_t  \left(1 - p_t - \sin^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\sqrt{2 \alpha \theta_0} \cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)} \right. \\
& \left.- \frac{1}{4}  \frac{\sqrt{2 \alpha \theta_0} \left(1 - 2 \cos^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)} \right]\dif t + \dif W_t \\
 = &\left[  \frac{  2  \dot{p}_t - \theta_t (1 - 2 p_t)  + (\alpha \theta_0 - \theta_t) \cos(- \sqrt{2 \alpha \theta_0 } Z_t) }{\sqrt{2 \alpha \theta_0} \sin{(- \sqrt{2 \alpha \theta_0} Z_t)}}  \right] \dif t + \dif W_t.  \label{eq:stindepSDE2}
\end{split}
\end{align}}

@@

Remark: Moreover, this obtained Gaussian distribution supports the validity of the choice of our model diffusion coefficient given by (\ref{eq:poly_diff}).

%{\color{red} Remind that Lamperti transform has the purpose tho check the consistency of the models, that is if the estimation of $\bm{\theta}$ with $V$ and $Z$ differs it is alarming.}

@@

Suppose that any of $M$ non-overlapping paths of the continuous-time It\^{o} process $V = \{ V_t, t  \in [0,T] \}$, each one starting at a different time $t_j$ with $j = 1, \dots, M$, is sampled at $N + 1$ equispaced discrete points with given length interval $\Delta$. Let $ V^{M,N + 1}=\left\{ V_{t_1^{N + 1}} , V_{t_2^{N + 1}} ,\ldots , V_{t_M^{N + 1}} \right\}$ denote this random sample, with $V_{t_j^{N + 1}} =\left\{ V_{t_j + i \Delta}\,, i = 0, \ldots, N \right\}$.

@@

However, closed-form solutions to initial-boundary value problems for time-inhomogeneous diffusions can be obtained only in a few cases (see, for example, \cite[Section 3.1]{eglix}). In our case, solving numerically (\ref{eq:fpk}) for the transition densities of the process $V$ at every transition step is computationally expensive. 
Several numerical techniques have been devised to obtain estimates for the unknown parameters of continuous-time SDE models with discrete observations (see, for example, \cite{prewo} for likelihood-based inference techniques, \cite{Sor} for an estimating function approach). As explained in the next subsection, we have considered approximate likelihood methods, similar in spirit to \cite[Section 11.4]{saso}.

@@

Gaussian approximations to the transition densities of nonlinear time-inhomogeneous SDEs are available through different algorithms \cite[Chapter 9]{saso}. However, as Figure \ref{fig:error_transitions} may suggest at first glance, the choice of a Gaussian density could be inadequate when straightly applied to approximate the transition density of the forecast error $V$ of the normalized wind power production.

@@

A suitable candidate for a surrogate transition density of $V$ is a Beta distribution on a compact interval parameterized by two positive shape parameters, $\xi_1, \xi_2$. {Recall that the choice of the beta proxy distribution is a natural choice as it is the invariant distribution of the Jacobi type processes.}

@@

The approximate log-likelihood $\tilde{\ell}(\cdot ; v^{M, N+1})$ of the observed sample $v^{M, N+1}$ can be expressed as
{\small
\begin{multline}
 \tilde{\ell} \left(\bm{\theta}; v^{M,N +1}\right) \\
 = \sum_{j=1}^M \sum_{i=1}^N \log  \left\{ \frac{1}{2(1 - \epsilon)} \frac{1}{B(\xi_1(t_{j,i}^-), \xi_2(t_{j,i}^-))} \left( \frac{v_{j,i} + 1 - \epsilon}{2(1 - \epsilon)} \right)^{\xi_1(t_{j,i}^-) -1}  \left( \frac{1 - \epsilon - v_{j,i}}{2(1 - \epsilon)} \right)^{\xi_2(t_{j,i}^-) -1} \right\},
\label{eq:loglikelihoodV}
\end{multline}
}
where the shape parameters $\xi_1(t_{j,i}^-)$ and $\xi_2(t_{j,i}^-)$, according to (\ref{param_transformed_beta}), depend on the limit quantities $\mu(t_{j,i}^-;\bm{\theta} )$ and $\sigma^2(t_{j,i}^-;\bm{\theta} )$ as $t\uparrow t_{j,i}$ that are computed solving numerically the initial-value problem (\ref{Vtmom}). $B(\xi_1,\xi_2)$ denotes the Beta distribution with parameters $\xi_1$ and $\xi_2$.

@@

{In this subsection, we aim to infer the model's parameters using optimization techniques. We start by finding an initial guess close enough to the optimal value, and from that point, start the optimization.}

@@

Our statistical analysis starts with partitioning the 147 segments of normalized wind power production, each 24-hours long. We select 73 non-contiguous segments for the models' calibration procedure, assigning them to the training set. The other 74 non-contiguous segments compose the test set. Such an allocation mechanism guarantees independence among the segments, matching the assumption we did in Section \ref{Section_5} to formulate the statistical models. {Additional cross-correlation tests were performed to ensure this assumption.}

@@

As an auxiliary verification, we plot the negative log-likelihood (negative version of (\ref{eq:loglikelihoodV})) as a function of the parameters, and we use additional minimization functions from MATLAB R2019b. {Moreover, we realized an additional inference utilizing the test sets to guarantee the robustness of our numerical methods.}

\begin{figure}[H]
\centering
\includegraphics[width=0.5\textwidth]{plots/LL/Log-Likelihood.eps}\includegraphics[width=0.5\textwidth]{plots/LL/Log-Likelihood_testing.eps}
\caption{{Negative log-likelihood's level sets for the training sets (plot on the left), and for the test sets (plot on the right). All optimal values are located over the curve $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively.}}
\label{fig:neg-LL}
\end{figure}
{On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood for both training and test sets.} The numerical values of each relevant point can be seen in Table (\ref{tab:optimal_values}). We set the optimal parameters in the $V$-space $(\theta_0^V,\alpha^V)=(1.93,0.050)$, as it is where the negative log-likelihood for the training sets reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{3}{c}{Training sets} & \multicolumn{3}{c}{Test sets} \\
 \midrule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$ & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.072 & 0.111 & 1.96 & 0.053 & 0.104 \\
 fminsearch & 1.14 & 0.076 & 0.097 & 1.64 & 0.054 & 0.089 \\
 fmincon & 1.58 & 0.062 & 0.097 & 1.63 & 0.055 & 0.089 \\
 fminunc & 1.54 & 0.063 & 0.097 & 1.96 & 0.045 & 0.089 \\
 Evaluations & 1.93 & 0.050 & 0.097 & 1.59 & 0.056 & 0.089 \\
 \bottomrule
\end{tabular}
\caption{Value points from Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
{We observe that all the local (possibly global) minimizers are located over the curves $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively. This effect shows that the optimization is more sensitive to the diffusion than to the drift.}\\
